Extended Data

### 2022-03-07 doc/manuscript/Dow_manuscript_Nature.Rmd
first draft of new intro content
@camerondow35 , FYI, and see what you think (when you have a chance). I still need to clean up/ make sure I've captured all the points.

#117

~~~~ (rework starting here)

@@

~~~~ (rework ending here)

@@

>>>>

~~~~ (new intro content starting here)

In temperate deciduous forests, spring warming generally lengthens the period over which trees have photosynthetically active leaves[@jeong_phenology_2011;@ibanez_forecasting_2010;@menzel_european_2006;@menzel_growing_1999;@lu_no_2022] and that over which the ecosystem is a net CO~2~ sink[@keenan_net_2014;@lu_no_2022].
Based on these observations, current terrestrial ecosystem models effectively assume that longer growing seasons lead to sustained increases in annual net CO~2~ uptake (i.e., net ecosystem exchange, _NEE_[@buermann_widespread_2018;@zani_increased_2020;@keenan_net_2014;@ahlstrom_robustness_2012].
However, model representation of C allocation to stem growth -- or woody aboveground net primary productivity ($ANPP_{woody}$) on the ecosystem level -- and its seasonality is simplistic relative to *known decoupling of photosynthate production and stem growth [@delpierre_temperate_2016;@delpierre_wood_2016]*. <!-- go back and check that these references are appropriate for this statement-->
Given that the C allocated to woody growth has the longest residence time in the ecosystem[@refs], understanding whether warming spring temperatures will lead to sustained increases in forest C sequestration requires understanding the sensitivity of $ANPP_{woody}$ to spring temperatures.

We have little evidence as to how stem growth and $ANPP_{woody}$ respond to warmer spring temperatures, particularly for temperate deciduous forests. 
Most studies on tree growth responses to warmer spring temperatures have focused on boreal or temperate conifers, which tend to respond to warmer spring temperatures with an earlier start to growth[@conifer_refs] and increased annual growth in mesic climates[@gao_earlier_2022;@more_conifer_refs]. 
Much less is known about the climate sensitivity of growth of deciduous broadleaf trees, although it has been observed that their growth is more dependent on conditions within the growing season then on growing season length[@etzold_number_2022]. 
Furthermore, recent evidence suggests that growth of broadleaf deciduous trees may be sink-saturated, such that longer growing seasons with more C fixation do not necessarily augment growth[@refs].
Tree-ring records, which can be used to examine relationships of annual growth to temperature but not to understand growth seasonality, reveal that growth of temperate deciduous trees tends to be most sensitive to temperature or potential evapotranspiration between late spring and early summer[@dorangeville_drought_2018;@helcoski_growing_2019], with some evidence that warmer springs may have a modest, but usually insignificant, positive effect on growth[@helcoski_growing_2019;@gao_earlier_2022].
These observations do not necessarily align with the finding that warming spring temperatures increase annual forest CO~2~ uptake in temperate deciduous forests[@keenan_net_2014].
Characterizing responses of stem growth to warming spring temperatures is critical to bridging this conceptual disconnect and understanding how forest biomass growth is likely to change as the climate warms.

~~~~ ( old version starting here)

@@

@@

~~~~ (end old version)
